Lineage Analysis of CRISPR and shRNA Screens

### 2018-08-24 lineages_18Q3.Rmd
updated Rmd/html w/ per-gene plots

```{r}
makeCRISPRlinplot <- function(g) {
  gene <- filter(crispr_data, Hugo_Symbol == g)
  sig <- filter(crispr_signif_lineage, Hugo_Symbol == g)
  gene <- merge(gene, sig, by = c("Hugo_Symbol", "lineage_name"))
  gene <- mutate(gene, lineage_name = reorder(lineage_name, p, mean))
  crispr_color <- as.character(gene$Color)
  names(crispr_color) <- gene$Mutation_Status
  score_range <- abs(range(gene$Score)[2] - range(gene$Score)[1])
  round_accuracy <- ifelse(score_range <= 2, 0.25,
                           ifelse(score_range <= 3, 0.5, 1.0))

  plot <- ggplot(data = gene, aes(x = lineage_name, y = Score)) +
    geom_point(alpha = 0.5, mapping = aes(color = Mutation_Status)) +
    scale_color_manual(values = crispr_color) +
    coord_cartesian(y = c(min(gene$Score), round_any(x = max(gene$Score), accuracy = round_accuracy, f = ceiling))) +
    geom_signif(data = gene, mapping = aes(xmin = lineage_name, xmax = lineage_name, annotations = paste(p.short, "\n", p.adj.short), y_position = max(gene$Score)), manual = TRUE, tip_length   = 0, size = 0, textsize = 3) +
    theme(legend.position = "top", axis.text.x = element_text(angle = 70, hjust = 1, size = 10)) +
    labs(y = "CERES Score",
         x = "Lineage",
         color = "Mutation Status",
         title = paste0(g, ": CERES score by cell line lineage"),
         subtitle = "Lineages sorted by increasing p-value; labels indicate unadjusted p-value / BH-corrected p-value.")
  return(plot)
}
```

@@

  pdf(paste0("./plots_18Q3/crispr_heatmap_all_c", clust_num, ".pdf"), width = 100, height = 100, paper = "special")
  draw(chm_all_plot, heatmap_legend_side = "right", annotation_legend_side = "right")
  seekViewport("annotation_Cell_Line_SSMD")
  grid.text("Cell Line SSMD", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Cas9_Activity")
  grid.text("Cas9 Activity", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Culture_Type")
  grid.text("Culture Type", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Primary_Tissue")
  grid.text("Primary Tissue", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Primary_Disease")
  grid.text("Primary Disease", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Gender")
  grid.text("Gender", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Source")
  grid.text("Source", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  dev.off()

@@

knitr::kable(crispr_signif[1:100, c(1, 5, 6, 7, 8, 10)], caption = "CRISPR Screen: Wilcoxon Test Results Comparing Mutant and Wildtype Cell Lines (Benjamini-Hochberg-corrected p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% footnote(general = "Wilcoxon test Benjamini-Hochberg-corrected p-values:\n*: p <= 0.05\n**: p <= 0.01\n***: p <= 0.001\n****: p <= 0.0001") %>% scroll_box(width = "900px", height = "450px")
```

@@

knitr::kable(shrna_signif[1:100, c(1, 5, 6, 7, 8, 10)], caption = "shRNA Screen: Wilcoxon Test Results Comparing Mutant and Wildtype Cell Lines (Benjamini-Hochberg-corrected p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% footnote(general = "Wilcoxon test Benjamini-Hochberg-corrected p-values:\n*: p <= 0.05\n**: p <= 0.01\n***: p <= 0.001\n****: p <= 0.0001") %>% scroll_box(width = "900px", height = "450px")
```

@@

Plot code:
```{r, eval=FALSE}
crispr_all_genes <- unique(crispr_data$Hugo_Symbol)
crispr_all_groups <- lapply(crispr_all_genes, makeCRISPRgrob)
crispr_all_groups_paths <- paste0(crispr_all_genes, "_crispr_grouped.png")
pwalk(list(crispr_all_groups_paths, crispr_all_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots", dpi = 300, width = 12, height = 12, units = "in")
```

Data management:
```{r, eval=FALSE}
crispr_all_bygene <- paste0(list.files("./plots_18Q3/crispr_grouped_plots", full.names = TRUE))
names(crispr_all_bygene) <- str_replace_all(crispr_all_bygene, c("_crispr_grouped.png" = "", "./plots_18Q3/crispr_grouped_plots/" = ""))
```

Plots:
```{r, eval=FALSE}
crispr_bygene_order <- c(intersect(crispr_signif$Hugo_Symbol, names(crispr_all_bygene)), setdiff(names(crispr_all_bygene), crispr_signif$Hugo_Symbol))
crispr_all_bygene <- crispr_all_bygene[match(crispr_bygene_order, names(crispr_all_bygene))]
bsselect(crispr_all_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

## Sorted lineage plots {.tabset .tabset-fade .tabset-pills}

### KRAS

```{r fig.height=10, fig.width=15}
makeCRISPRlinplot("KRAS")
```

### TP53

```{r fig.height=10, fig.width=15}
makeCRISPRlinplot("TP53")
```

### BRAF

```{r fig.height=10, fig.width=15}
makeCRISPRlinplot("BRAF")
```

### NRAS

```{r fig.height=10, fig.width=15}
makeCRISPRlinplot("NRAS")
```

>>>>

```{r}
makeCRISPRlinplot <- function(g) {
  sig <- filter(crispr_signif_lineage, Hugo_Symbol == g)
  if(nrow(sig) == 0) {
    return(NULL)
  }
  else {
    gene <- filter(crispr_data, Hugo_Symbol == g)
    gene <- merge(gene, sig, by = c("Hugo_Symbol", "lineage_name"))
    gene <- mutate(gene, lineage_name = reorder(lineage_name, p, mean))
    crispr_color <- as.character(gene$Color)
    names(crispr_color) <- gene$Mutation_Status
    score_range <- abs(range(gene$Score)[2] - range(gene$Score)[1])
    round_accuracy <- ifelse(score_range <= 2, 0.25,
                             ifelse(score_range <= 3, 0.5, 1.0))

    plot <- ggplot(data = gene, aes(x = lineage_name, y = Score)) +
      geom_point(alpha = 0.5, mapping = aes(color = Mutation_Status)) +
      scale_color_manual(values = crispr_color) +
      coord_cartesian(y = c(min(gene$Score), round_any(x = max(gene$Score), accuracy = round_accuracy, f = ceiling))) +
      geom_signif(data = gene, mapping = aes(xmin = lineage_name, xmax = lineage_name, annotations = paste(p.short, "\n", p.adj.short), y_position = max(gene$Score)), manual = TRUE, tip_length   = 0, size = 0, textsize = 3) +
      theme(legend.position = "top", axis.text.x = element_text(angle = 70, hjust = 1, size = 10)) +
      labs(y = "CERES Score",
           x = "Lineage",
           color = "Mutation Status",
           title = paste0(g, ": CERES score by cell line lineage"),
           subtitle = "Lineages sorted by increasing p-value; labels indicate unadjusted p-value / BH-corrected p-value.")
    return(plot)
  }
}
```

@@

  pdf(paste0("./plots_18Q3/crispr_heatmap_clustered/crispr_heatmap_all_c", clust_num, ".pdf"), width = 100, height = 100, paper = "special")
  draw(chm_all_plot, heatmap_legend_side = "right", annotation_legend_side = "right")
  seekViewport("annotation_Cell_Line_SSMD")
  grid.text("Cell Line SSMD", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Cas9_Activity")
  grid.text("Cas9 Activity", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Culture_Type")
  grid.text("Culture Type", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Primary_Tissue")
  grid.text("Primary Tissue", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Primary_Disease")
  grid.text("Primary Disease", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Gender")
  grid.text("Gender", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  seekViewport("annotation_Source")
  grid.text("Source", unit(1, "npc") + unit(2, "mm"), 0.5, gp = gpar(fontsize = 30), default.units = "npc", just = "left")
  dev.off()

@@

knitr::kable(filter(crispr_signif, p < 0.1)[, c(1, 5, 6, 7, 8, 10)], caption = "CRISPR Screen: Wilcoxon Test Results Comparing Mutant and Wildtype Cell Lines, p < 0.1 (Benjamini-Hochberg-corrected p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% footnote(general = "Wilcoxon test Benjamini-Hochberg-corrected p-values:\n*: p <= 0.05\n**: p <= 0.01\n***: p <= 0.001\n****: p <= 0.0001") %>% scroll_box(width = "900px", height = "450px")
```

@@

knitr::kable(filter(shrna_signif, p < 0.1)[, c(1, 5, 6, 7, 8, 10)], caption = "shRNA Screen: Wilcoxon Test Results Comparing Mutant and Wildtype Cell Lines, p < 0.1  (Benjamini-Hochberg-corrected p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% footnote(general = "Wilcoxon test Benjamini-Hochberg-corrected p-values:\n*: p <= 0.05\n**: p <= 0.01\n***: p <= 0.001\n****: p <= 0.0001") %>% scroll_box(width = "900px", height = "450px")
```

@@

Plot code:
```{r, eval=FALSE}
crispr_data_cgc <- filter(crispr_data, Hugo_Symbol %in% cgc$Hugo_Symbol)
crispr_all_genes <- unique(crispr_data_cgc$Hugo_Symbol)
crispr_all_groups <- lapply(crispr_all_genes, makeCRISPRgrob)
crispr_all_groups_paths <- paste0(crispr_all_genes, "_crispr_grouped.png")
pwalk(list(crispr_all_groups_paths, crispr_all_groups), ggsave, path = "./plots_18Q3/crispr_grouped_plots_cgc", dpi = 300, width = 12, height = 12, units = "in")
```

Data management:
```{r}
crispr_all_bygene <- paste0(list.files("./plots_18Q3/crispr_grouped_plots_cgc", full.names = TRUE))
names(crispr_all_bygene) <- str_replace_all(crispr_all_bygene, c("_crispr_grouped.png" = "", "./plots_18Q3/crispr_grouped_plots_cgc/" = ""))
crispr_bygene_order <- c(intersect(crispr_signif$Hugo_Symbol, names(crispr_all_bygene)), setdiff(names(crispr_all_bygene), crispr_signif$Hugo_Symbol))
crispr_all_bygene <- crispr_all_bygene[match(crispr_bygene_order, names(crispr_all_bygene))]
bsselect(crispr_all_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

## Sorted lineage plots

Plot code:
```{r, eval=FALSE}
crispr_data_cgc <- filter(crispr_data, Hugo_Symbol %in% cgc$Hugo_Symbol)
crispr_all_genes <- unique(crispr_data_cgc$Hugo_Symbol)
crispr_all_lins <- lapply(crispr_all_genes, makeCRISPRlinplot)
crispr_all_lins_paths <- paste0(crispr_all_genes, "_crispr_lineage.png")
pwalk(list(crispr_all_lins_paths, crispr_all_lins), ggsave, path = "./plots_18Q3/crispr_lineage_plots_cgc", dpi = 300, width = 12, height = 12, units = "in")
```

```{r}
makeCRISPRlinplot("HMGA1")
test <- filter(crispr_signif_lineage, Hugo_Symbol == "HMGA1")
```

Data management:
```{r}
crispr_all_bygene <- paste0(list.files("./plots_18Q3/crispr_lineage_plots_cgc", full.names = TRUE))
names(crispr_all_bygene) <- str_replace_all(crispr_all_bygene, c("_crispr_lineage.png" = "", "./plots_18Q3/crispr_lineage_plots_cgc/" = ""))
crispr_bygene_order <- c(intersect(crispr_signif$Hugo_Symbol, names(crispr_all_bygene)), setdiff(names(crispr_all_bygene), crispr_signif$Hugo_Symbol))
crispr_all_bygene <- crispr_all_bygene[match(crispr_bygene_order, names(crispr_all_bygene))]
bsselect(crispr_all_bygene, type = "img", selected = "KRAS", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```
